---
title: "Data 621 Homework 1"
author: "Group 2"
date: "September 21, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, warning=FALSE, message=FALSE}
library(knitr)
library(kableExtra)
library(psych)
library(DT)
library(ggplot2)
library(reshape2)
library(gridExtra)
#library(psychometric)
library(corrplot)
#library(ggpubr)
#library(matlib)
#library(matrixcalc)
```

---

## Group Members

This is a public page. I'm not sure if people want their names listed here. 
```{r echo=FALSE}
#- Yohannes Deboch
#- Benjamin Horvath
#- Santosh Manjrekar
#- Samriti Malhotra 
#- Sherranette Tinapunan
```

---

## Overview

In this homework assignment, you will explore, analyze and model a data set containing approximately 2200 records. Each record represents a professional baseball team from the years <b>1871 to 2006</b> inclusive. Each record has the performance of the team for the given year, with all of the statistics <b>adjusted to match the performance of a 162 game season</b>.

Your objective is to build a multiple linear regression model on the training data to predict the number of wins
for the team.

---

## Training Data Set


The training data has <b>17</b> columns and <b>2,276 rows</b>. <br/>

The explanatory columns are broken down into <b>four categories</b>: <br/>

- Batting
- Base run 
- Pitching
- Fielding

Below you will see a preview of the columns and the first few observations broken down into these four categories. 

<br/>

```{r echo=FALSE}
data <- read.csv(file="moneyball-training-data.csv", header=TRUE, sep=",")
```

---

## Response Variable 

### <span style="color:blue">TARGET_WINS</span>

The variable <b>TARGET_WINS</b> is the <b>number of wins</b> of a professional baseball team for a given year. The year is not part of the data set.
This is the <b>dependent</b> variable that our models will attempt to predict. 

<br/> 
                   
```{r echo=FALSE}
datatable(head(data[c(2)]), options=list(searching = FALSE, lengthChange = FALSE, info=FALSE, dom='t', autoWidth = TRUE), width=200)
```
Above is first few rows of dependent variable. 

<br/> 

#### DESCRIPTIVE STATISTCS 

```{r echo=FALSE}
datatable(data.frame(unclass(summary(data[c(2)])), check.names = FALSE, stringsAsFactors = TRUE), rownames=FALSE, width=200, options=list(dom='t'))
```

```{r echo=FALSE}
datatable(psych::describe(data[c(1:2)], skew=FALSE)[2,c(2,4,8)], options = list(dom='t'), width=200) %>% formatRound(c("sd", "se"), 2)
```

<br/> 

#### PLOTS 

```{r echo=FALSE, fig.wdith=8.5}
par(mfrow=c(1,1))
p1 <- ggplot(data, aes(x=TARGET_WINS)) + geom_histogram(binwidth=5) + geom_vline(xintercept=mean(data$TARGET_WINS)) + theme(axis.title.x=element_blank()) 
p2 <- ggplot(data, aes(x=TARGET_WINS)) + stat_density() + theme(axis.title.x=element_blank()) 
grid.arrange(p1,p2, ncol=2, top="TARGET_WINS")
boxplot(data[c(2)])
```

As you can see, the distribution of the number of wins is unimodal and skewed to the left with some outliers towards the tail. It looks approximately normal. The minimum number of wins for a team is 0 and the maximum is 146. The mean is 80.79. 

The boxplot above shows that there are suspected outliers at both ends. 

```{r echo=FALSE}
summary(data$TARGET_WINS)
```

---

<br/>

## Explanatory Variables 

### <span style="color:blue">1. BATTING VARIABLES (7)</span>

<br /> 

#### DESCRIPTION AND THEORETICAL EFFECT


| VARIABLE NAME    | DEFINITION                              | THEORETICAL EFFECT             |
|------------------|-----------------------------------------|--------------------------------|
| TEAM_BATTING_H   | Base Hits by batters (1B,2B,3B,HR)      | Positive Impact on Wins        |
| TEAM_BATTING_2B  | Doubles by batters (2B)                 | Positive Impact on Wins        |
| TEAM_BATTING_3B  | Triples by batters (3B)                 | Positive Impact on Wins        |
| TEAM_BATTING_HR  | Homeruns by batters (4B)                | Positive Impact on Wins        |
| TEAM_BATTING_BB  | Walks by batters                        | Positive Impact on Wins        |
| TEAM_BATTING_HBP |  Batters hit by pitch (get a free base) | Positive Impact on Wins        |
| TEAM_BATTING_SO  | Strikeouts by batters                   | <b>Negative Impact on Wins</b> |

<br/> 
                   
```{r echo=FALSE}
datatable(head(data[c(3:8, 11)]), options = list(searching = FALSE, lengthChange = FALSE, info=FALSE, dom='t'))
```
Above is view of first few rows of batting variables. 

<br/> 

#### DESCRIPTIVE STATISTCS 

As you can see, two variables have some N/A values (n < 2276). Particularly, <b>TEAM_BATTING_HPB only has 191 values that are not missing</b>. 

```{r echo=FALSE}
datatable(data.frame(unclass(summary(data[c(3:8, 11)])), check.names = FALSE, stringsAsFactors = TRUE), rownames=FALSE, options = list(searching = FALSE, lengthChange = FALSE, dom='t'))
```

```{r echo=FALSE}
datatable(psych::describe(data[c(3:8, 11)], skew=FALSE)[c(2,4,8)], options = list(dom='t'), width=200) %>% formatRound(c("sd", "se"), 2)
```

<br />

#### PLOTS

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=8.5}
melt.data <- melt(data[c(3:8,11)])
ggplot(data = melt.data, aes(x = value)) + stat_density() + facet_wrap(~variable, scales = "free")
```

<br />

```{r echo=FALSE, warning=FALSE, message=FALSE, fig.width=8.5}
par(mfrow=c(1,3))
boxplot(data[c(3)], main=colnames(data)[3])
boxplot(data[c(4)], main=colnames(data)[4])
boxplot(data[c(5)], main=colnames(data)[5])
boxplot(data[c(6)], main=colnames(data)[6])
boxplot(data[c(7)], main=colnames(data)[7])
boxplot(data[c(8)], main=colnames(data)[8])
boxplot(data[c(11)], main=colnames(data)[11])
```

#### CORRELATION OF BATTING VARIABLES

See <i>Pairings of all Variables </i> to view correlation table and scatter plots of all variables. 

```{r echo=FALSE}
cor <- cor(data[c(2:8,11)], method = "pearson", use = "complete.obs") 
datatable(cor, options=list(dom="t")) %>% formatRound(c(1:8), 2)
```

```{r echo=FALSE, fig.width=20, fig.height=20}
#pdf("pairs_output.pdf")
pairs(data[c(2:8,11)]) #Batting
#dev.off()
```


<br/> 

### <span style="color:blue">2. BASE RUN VARIABLES (2)</span>

<br/> 

#### DESCRIPTION AND THEORETICAL EFFECT

| VARIABLE NAME   | DEFINITION      | THEORETICAL EFFECT           |
|-----------------|-----------------|------------------------------|
|TEAM_BASERUN_SB  |Stolen bases     |Positive Impact on Wins       |
|TEAM_BASERUN_CS  |Caught stealing  |<b>Negative Impact on Wins</b>|

<br/> 

```{r echo=FALSE}
datatable(head(data[c(9:10)]), options = list(searching = FALSE, lengthChange = FALSE, dom='t'),  width=200)
```
Above is view of first few rows of base run variables. 

<br/> 

#### DESCRIPTIVE STATISTCS 

As you can see, both variables have some N/A values (n < 2276). 

```{r echo=FALSE}
datatable(data.frame(unclass(summary(data[c(9:10)])), check.names = FALSE, stringsAsFactors = TRUE), rownames=FALSE, options = list(searching = FALSE, lengthChange = FALSE, dom='t'), width=200)
```

```{r echo=FALSE}
datatable(psych::describe(data[c(9:10)], skew=FALSE)[c(2,4,8)], options = list(dom='t'),  width=200) %>% formatRound(c("sd", "se"), 2)
```

<br/> 

#### PLOTS

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=6}
melt.data <- melt(data[c(9:10)])
ggplot(data = melt.data, aes(x = value)) + stat_density() + facet_wrap(~variable, scales = "free")
```

<br /> 

```{r echo=FALSE, warning=FALSE, message=FALSE, fig.width=6}
par(mfrow=c(1,2))
boxplot(data[c(9)], main=colnames(data)[9])
boxplot(data[c(10)], main=colnames(data)[10])
```


#### CORRELATION OF BASE RUN VARIABLES

See <i>Pairings of all Variables </i> to view correlation table and scatter plots of all variables. 

```{r echo=FALSE}
cor <- cor(data[c(2, 9:10)], method = "pearson", use = "complete.obs") 
datatable(cor, options=list(dom="t")) %>% formatRound(c(1:8), 2)
```

```{r echo=FALSE, fig.width=5, fig.height=5}
#pdf("pairs_output.pdf")
pairs(data[c(2, 9:10)]) #Batting
#dev.off()
```

### <span style="color:blue">3. PITCHING VARIABLES (4)</span>

<br /> 

#### DESCRIPTION AND THEORETICAL EFFECT

| VARIABLE NAME   | DEFINITION             | THEORETICAL EFFECT            |
|-----------------|------------------------|-------------------------------|
|TEAM_PITCHING_BB | Walks allowed          |<b>Negative Impact on Wins</b> |
|TEAM_PITCHING_H  | Hits allowed           |<b>Negative Impact on Wins</b> |
|TEAM_PITCHING_HR | Homeruns allowed       |<b>Negative Impact on Wins</b> |
|TEAM_PITCHING_SO | Strikeouts by pitchers | Positive Impact on Wins       |


```{r echo=FALSE}
datatable(head(data[c(12:15)]), options = list(searching = FALSE, lengthChange = FALSE, dom='t'), width=400)
```
Above is view of first few rows of pitching variables. 

<br/> 

#### DESCRIPTIVE STATISTCS 

As you can see, one variable has some N/A values (n < 2276). 

```{r echo=FALSE}
datatable(data.frame(unclass(summary(data[c(12:15)])), check.names = FALSE, stringsAsFactors = TRUE), rownames=FALSE, options = list(searching = FALSE, lengthChange = FALSE, dom='t'), width=400)
```

```{r echo=FALSE}
datatable(psych::describe(data[c(12:15)], skew=FALSE)[c(2,4,8)], options = list(dom='t'), width=400) %>% formatRound(c("sd", "se"), 2)
```

<br/>

#### PLOTS 

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=8}
melt.data <- melt(data[c(12:15)])
ggplot(data = melt.data, aes(x = value)) + stat_density() + facet_wrap(~variable, scales = "free")
```

<br /> 

```{r echo=FALSE, warning=FALSE, message=FALSE, fig.width=6}
par(mfrow=c(1,2))
boxplot(data[c(12)], main=colnames(data)[12])
boxplot(data[c(13)], main=colnames(data)[13])
boxplot(data[c(14)], main=colnames(data)[14])
boxplot(data[c(15)], main=colnames(data)[15])
```

#### CORRELATION OF PITCHING VARIABLES 

See <i>Pairings of all Variables </i> to view correlation table and scatter plots of all variables. 

```{r echo=FALSE}
cor <- cor(data[c(2, 12:15)], method = "pearson", use = "complete.obs") 
datatable(cor, options=list(dom="t")) %>% formatRound(c(1:8), 2)
```

```{r echo=FALSE, fig.width=7, fig.height=7}
#pdf("pairs_output.pdf")
pairs(data[c(2, 12:15)]) #Batting
#dev.off()
```

### <span style="color:blue">4. FIELDING VARIABLES (2)</span>

<br /> 

#### DESCRIPTION AND THEORETICAL EFFECT

| VARIABLE NAME   | DEFINITION  | THEORETICAL EFFECT            |
|-----------------|-------------|-------------------------------|
|TEAM_FIELDING_E  |Errors       | <b>Negative Impact on Wins</b>|
|TEAM_FIELDING_DP |Double Plays | Positive Impact on Wins       |


```{r echo=FALSE}
datatable(head(data[c(16:17)]), options = list(searching = FALSE, lengthChange = FALSE, dom='t'), width=200)
```
Above is view of first few rows of fielding variables. 

<br/> 

#### DESCRIPTIVE STATISTCS 

As you can see, one variable has some N/A values (n < 2276)

```{r echo=FALSE}
datatable(data.frame(unclass(summary(data[c(16:17)])), check.names = FALSE, stringsAsFactors = TRUE), rownames=FALSE, options = list(searching = FALSE, lengthChange = FALSE, dom='t'), width=200)
```

```{r echo=FALSE}
datatable(psych::describe(data[c(16:17)], skew=FALSE)[c(2,4,8)], options = list(dom='t'), width=200) %>% formatRound(c("sd", "se"), 2)
```

<br /> 

#### PLOTS 

```{r echo=FALSE, message=FALSE, warning=FALSE}
melt.data <- melt(data[c(16:17)])
ggplot(data = melt.data, aes(x = value)) + stat_density() + facet_wrap(~variable, scales = "free")
```

<br /> 

```{r echo=FALSE, warning=FALSE, message=FALSE, fig.width=6}
par(mfrow=c(1,2))
boxplot(data[c(16)], main=colnames(data)[16])
boxplot(data[c(17)], main=colnames(data)[17])
```

#### CORRELATION OF FIELDING VARIABLES 

See <i>Pairings of all Variables </i> to view correlation table and scatter plots of all variables. 

```{r echo=FALSE}
cor <- cor(data[c(2, 16:17)], method = "pearson", use = "complete.obs") 
datatable(cor, options=list(dom="t"), width=200) %>% formatRound(c(1:8), 2)
```

```{r echo=FALSE, fig.width=5, fig.height=5}
#pdf("pairs_output.pdf")
pairs(data[c(2, 16:17)]) #Batting
#dev.off()
```

---

### Pairings of all Variables 

#### CORRELATION OF ALL VARIABLES

```{r echo=FALSE}
cor <- cor(data[c(2:17)], method = "pearson", use = "complete.obs")
datatable(cor[,c(1, 2:7,10)], options=list(dom="t"), width=400) %>% formatRound(c(1:8), 2)
datatable(cor[,c(1,8:9)], options=list(dom="t"), width=400) %>% formatRound(c(1:3), 2)
datatable(cor[,c(1,11:14)], options=list(dom="t"), width=400) %>% formatRound(c(1:5), 2)
datatable(cor[,c(1,15:16)], options=list(dom="t"), width=400) %>% formatRound(c(1:3), 2)
```

#### SCATTER PLOTS OF ALL VARIABLES 

```{r echo=FALSE, fig.width=20, fig.height=20}
#pdf("pairs_output.pdf")
pairs(data[c(2:17)]) #Batting
#dev.off()
```


---

<br/>

## MISSING DATA

There are 6 explanatory variables with missing values. 

Results of simple linear regression of these variables with missing data (listed below) suggest that predictors are not significant at the 5% level. 
See section <i>Simple Linear Regression of each Variable</i> for more details. 
<b>Recommendation is to drop these from the models.</b>


- TEAM_BATTING_HBP - 2085 missing values (drop)
- TEAM_BATTING_SO - 102 missing values (drop)
- TEAM_BASERUN_CS - 772 missing (drop)
- TEAM_FIELDING_DP - 286 missing (drop)


Results of simple linear regression of variables (on TARGET_WINS) listed below suggest that predictors are significant.


- TEAM_BASERUN_SB - 131 missing values (ignore observations with missing data)
- TEAM_PITCHING_SO - 102 missing values (ignore observations with missing data)

---

<br /> 

# Simple Linear Regressions of each Variable

This section investigates more closely the relationship of each explanatory variable with the response variable TARGET_WINS. This would also give us some some insights with how to handle some of the variables with missing values. 


## BATTING VARIABLES

### TEAM_BATTING_H

Base hits by batter (1B, 2B, 3B, HR).
This variable is the sum of 1B, 2B, 3B, and HR. 

No Missing values. 

Positive relationship (expected theoretical effect is positive). No obvious curved patterns observed from visually inspecting the scatter plots only.

Linear regression predictor coefficient is 0.042353 (contributes 0.04 points to TARGET_WINS with every unit increase). The p-value is < 2e-16 *** (predictor is significant). The R-squared is 0.1511 (explains about 15% of variability of response variable).

```{r echo=FALSE, warning=FALSE}
ggplot(data, aes(x=data$TEAM_BATTING_H, y=data$TARGET_WINS)) + geom_point() + geom_smooth(method=lm) + 
  labs(x="TEAM_BATTING_H", y="TARGET_WINS", title="")
```

```{r}
summary(lm(TARGET_WINS ~ TEAM_BATTING_H, data=data))
```

---

### TEAM_BATTING_2B

Doubles by batters (2B).

No Missing values. 

Positive relationship (expected theoretical effect is positive). No obvious curved patterns observed from visually inspecting the scatter plots only.

Linear regression predictor coefficient is 0.097305 (points contributed to TARGET_WINS with every unit increase). The p-value is <2e-16 *** (predictor is significant). The R-squared is 0.08358 (explains about 8% of variability of response variable).


```{r echo=FALSE, warning=FALSE}
ggplot(data, aes(x=data$TEAM_BATTING_2B, y=data$TARGET_WINS)) + geom_point() + geom_smooth(method=lm) +
  labs(x="TEAM_BATTING_2B", y="TARGET_WINS", title="")
```

```{r}
summary(lm(TARGET_WINS ~ TEAM_BATTING_2B, data=data))
```

---

### TEAM_BATTING_3B

Tripples by batter (3B).

No missing values. 

Positive relationship (expected theoretical effect is positive). No obvious curved patterns observed from visually inspecting the scatter plots only.

Linear regression predictor coefficient is 0.0804 (points contributed to TARGET_WINS with every unit increase). The p-value is 8.22e-12 *** (predictor is significant). The R-squared is  0.02034 (explains about 2% of variability of response variable).


```{r echo=FALSE, warning=FALSE}
ggplot(data, aes(x=data$TEAM_BATTING_3B, y=data$TARGET_WINS)) + geom_point() + geom_smooth(method=lm) + labs(x="TEAM_BATTING_3B", y="TARGET_WINS", title="")
```

```{r}
summary(lm(TARGET_WINS ~ TEAM_BATTING_3B, data=data))
```

---

### TEAM_BATTING_HR

Homeruns by batter (4B).

No missing values. 

Positive relationship (expected theoretical effect is positive). No obvious curved patterns observed from visually inspecting the scatter plots only.

Linear regression predictor coefficient is 0.04583 (points contributed to TARGET_WINS with every unit increase). The p-value is <2e-16 *** (predictor is significant). The R-squared is 0.03103 (explains about 3% of variability of response variable).


```{r echo=FALSE, warning=FALSE}
ggplot(data, aes(x=data$TEAM_BATTING_HR, y=data$TARGET_WINS)) + geom_point() + geom_smooth(method=lm) + 
  labs(x="TEAM_BATTING_HR", y="TARGET_WINS", title="")
```

---

### TEAM_BATTING_BB

Walks by batter. 

No missing values.

Positive relationship (expected theoretical effect is positive). No obvious curved patterns observed from visually inspecting the scatter plots only.

Linear regression predictor coefficient is 0.029863 (points contributed to TARGET_WINS with every unit increase). The p-value is <2e-16 *** (predictor is significant). The R-squared is 0.05408 (explains about 5% of variability of response variable).


```{r echo=FALSE, warning=FALSE}
ggplot(data, aes(x=data$TEAM_BATTING_BB, y=data$TARGET_WINS)) + geom_point() + geom_smooth(method=lm) + 
  labs(x="TEAM_BATTING_BB", y="TARGET_WINS", title="")
```

```{r}
summary(lm(TARGET_WINS ~ TEAM_BATTING_BB, data=data))
```

---

### TEAM_BATTING_SO (102 Missing)

Strikeouts by batter.

102 missing values. 

Simple linear regression result below suggests that this variable is not significant. 

The recommendation is to drop this variable from the model. 

Negative relationship (expected theoretical effect is negative). No obvious curved patterns observed from visually inspecting the scatter plots only.

Linear regression predictor coefficient is -0.001990 (points contributed to TARGET_WINS with every unit increase). The p-value is 0.139 (predictor is not significant at 5% level). The R-squared is 0.001008 (explains less than 1% of variability of response variable).


```{r echo=FALSE, warning=FALSE}
ggplot(data, aes(x=data$TEAM_BATTING_SO, y=data$TARGET_WINS)) + geom_point() + geom_smooth(method=lm) + 
  labs(x="TEAM_BATTING_SO", y="TARGET_WINS", title="")
```

```{r}
summary(lm(TARGET_WINS ~ TEAM_BATTING_SO, data=data))
```


---

### TEAM_BATTING_HBP (2085 Missing)

Batters hit by pitch. 

2085 missing values. 

Simple linear regression result suggests that this variable is not significant. 

Recommendation is to drop this variable from the model. 

Positive relationship (expected theoretical effect is positive). No obvious curved patterns observed from visually inspecting the scatter plots only.

Linear regression predictor coefficient is  0.06867 (points contributed to TARGET_WINS with every unit increase). The p-value is  0.312 (predictor is not significant at 5% level). The R-squared is 0.005403 (explains less than 1% of variability of response variable).


```{r echo=FALSE, warning=FALSE}
ggplot(data, aes(x=data$TEAM_BATTING_HBP, y=data$TARGET_WINS)) + geom_point() + geom_smooth(method=lm) + 
  labs(x="TEAM_BATTING_HBP", y="TARGET_WINS", title="")
```

```{r}
summary(lm(TARGET_WINS ~ TEAM_BATTING_HBP, data=data))
```

---

## PITCHING VARIABLES


### TEAM_PITCHING_BB

Walks allowed. 

No missing values. 

Positive relationship (expected theoretical impact <b>negative</b>). No obvious curved patterns observed from visually inspecting the scatter plots only. Noticeable outliers that may have strong influence on the linear regression line. Most points on the scatter plot are below 1000. A second scatter plot of points under 1000 continues to show a positive relationship. 

All Points:
Linear regression predictor coefficient is 0.01176 (points contributed to TARGET_WINS with every unit increase). The p-value is2.78e-09 *** (predictor is significant). The R-squared is 0.01542 (explains about 1.5% of variability of response variable).

Under 1000:
Linear regression predictor coefficient is 0.028531 (points contributed to TARGET_WINS with every unit increase). The p-value is <2e-16 *** (predictor is significant). The R-squared is 0.04401 (explains about 4.4% of variability of response variable).


```{r echo=FALSE, warning=FALSE}
ggplot(data, aes(x=data$TEAM_PITCHING_BB, y=data$TARGET_WINS)) + geom_point() + geom_smooth(method=lm) + 
  labs(x="TEAM_PITCHING_BB", y="TARGET_WINS", title="")
```

```{r}
summary(lm(TARGET_WINS ~ TEAM_PITCHING_BB, data=data))
```

Points under 1000.

```{r echo=FALSE, warning=FALSE}
ggplot(data[data$TEAM_PITCHING_BB<1000,], aes(x=data$TEAM_PITCHING_BB[data$TEAM_PITCHING_BB<1000], y=data$TARGET_WINS[data$TEAM_PITCHING_BB<1000])) + geom_point() + geom_smooth(method=lm) + labs(x="TEAM_PITCHING_BB", y="TARGET_WINS", title="")
```

```{r}
summary(lm(TARGET_WINS ~ TEAM_PITCHING_BB, data=data[data$TEAM_PITCHING_BB<1000,]))
```

---

### TEAM_PITCHING_H

Hits allowed. 

No missing values. 

Negative relationship (expected theoretical effect is negative). No obvious curved patterns observed from visually inspecting the scatter plots only. Most points on the scatter plot are below 5000. A second scatter plot that only looks at points below 5000 suggest that the relationship is <b>positive</b>. 

All points: 
Linear regression predictor coefficient is -0.0012309 (points contributed to TARGET_WINS with every unit increase). The p-value is 1.46e-07 *** (predictor is significant). The R-squared is 0.01209 (explains about 1.2% of variability of response variable).

Points under 5000:
Linear regression predictor coefficient is 0.003825 (points contributed to TARGET_WINS with every unit increase). The p-value is 3.07e-07 *** (predictor is significant). The R-squared is 0.01167 (explains about 1.2% of variability of response variable).


```{r echo=FALSE, warning=FALSE}
ggplot(data, aes(x=data$TEAM_PITCHING_H, y=data$TARGET_WINS)) + geom_point() + geom_smooth(method=lm) + 
  labs(x="TEAM_PITCHING_H", y="TARGET_WINS", title="")
```

```{r}
summary(lm(TARGET_WINS ~ TEAM_PITCHING_H, data=data))
```

Points under 5000. 

```{r echo=FALSE}
ggplot(data[data$TEAM_PITCHING_H<5000,], aes(x=data$TEAM_PITCHING_H[data$TEAM_PITCHING_H<5000], y=data$TARGET_WINS[data$TEAM_PITCHING_H<5000])) + 
  geom_point() + geom_smooth(method=lm) + labs(x="TEAM_PITCHING_H", y="TARGET_WINS", title="")
```

```{r}
summary(lm(TARGET_WINS ~ TEAM_PITCHING_H, data=data[data$TEAM_PITCHING_H<5000,]))
```

---

### TEAM_PITCHING_HR

Home runs allowed.

No missing values. 

Positive relationship (expected theoretical effect is negative). No obvious curved patterns observed from visually inspecting the scatter plots only.

Linear regression predictor coefficient is 0.048572  (points contributed to TARGET_WINS with every unit increase). The p-value is <2e-16 *** (predictor is significant). The R-squared is 0.03573 (explains about 3.6% of variability of response variable).


```{r echo=FALSE, warning=FALSE}
ggplot(data, aes(x=data$TEAM_PITCHING_HR, y=data$TARGET_WINS)) + geom_point() + geom_smooth(method=lm) + 
  labs(x="TEAM_PITCHING_HR", y="TARGET_WINS", title="")
```

```{r}
summary(lm(TARGET_WINS ~ TEAM_PITCHING_HR, data=data))
```

---

### TEAM_PITCHING_SO (102 Missing)

Strikeouts by pitcher. 

102 missing values.

Simple linear regression result suggests that this variable is significant.

Negative relationship (expected theoretical effect is positive). No obvious curved patterns observed from visually inspecting the scatter plots only. Outliers may have a strong influence on the regression line. Most points are under 2000. A second scatter plot that only looks at points below 2000 also shows a negative relationship. 

Linear regression predictor coefficient is -0.0022085  (points contributed to TARGET_WINS with every unit increase). The p-value is 0.000252 *** (predictor is significant). The R-squared is 0.006152 (explains less than 1% of variability of response variable).


```{r echo=FALSE, warning=FALSE}
ggplot(data, aes(x=data$TEAM_PITCHING_SO, y=data$TARGET_WINS)) + 
  geom_point() + geom_smooth(method=lm) + labs(x="TEAM_PITCHING_SO", y="TARGET_WINS", title="")
```

```{r}
summary(lm(TARGET_WINS ~ TEAM_PITCHING_SO, data=data))
```

Scatter plot of points below 2000.

```{r echo=FALSE, warning=FALSE}
ggplot(data[data$TEAM_PITCHING_SO<2000,], aes(x=data$TEAM_PITCHING_SO[data$TEAM_PITCHING_SO<2000], y=data$TARGET_WINS[data$TEAM_PITCHING_SO<2000])) + 
  geom_point() + geom_smooth(method=lm) + labs(x="TEAM_PITCHING_SO", y="TARGET_WINS", title="")
```

```{r}
summary(lm(TARGET_WINS ~ TEAM_PITCHING_SO, data=data[data$TEAM_PITCHING_SO<2000,]))
```

---

## BASE RUN VARIABLES


### TEAM_BASERUN_SB (131 Missing)

Stolen bases. 

131 missing values. 

Results of simple linear regression below suggests that this variable is significant.  

Positive relationship (expected theoretical effect is positive). No obvious curved patterns observed from visually inspecting the scatter plots only.

Linear regression predictor coefficient is 0.02273 (points contributed to TARGET_WINS with every unit increase). The p-value is 3.3e-10 *** (predictor is significant). The R-squared is 0.01826 (explains about 1.8% of variability of response variable).


```{r echo=FALSE, warning=FALSE}
ggplot(data, aes(x=data$TEAM_BASERUN_SB, y=data$TARGET_WINS)) + 
  geom_point() + geom_smooth(method=lm) + labs(x="TEAM_BASERUN_SB", y="TARGET_WINS", title="")
```

```{r}
summary(lm(TARGET_WINS ~ TEAM_BASERUN_SB, data=data))
```

---

### TEAM_BASERUN_CS (772 Missing)

Caught stealing. 

772 missing values. Results of simple linear regression below suggests that this variable is not significant. 

Recommendation is to drop this variable from the model. 

Positive relationship. No obvious curved patterns observed from visually inspecting the scatter plots only.

Linear regression predictor coefficient is 0.01314 (points contributed to TARGET_WINS with every unit increase). The p-value is 0.385 (predictor is not significant at the 5% level). The R-squared is 0.0005019 (explains less than 1% variability of response variable).


```{r echo=FALSE, warning=FALSE}
ggplot(data, aes(x=data$TEAM_BASERUN_CS, y=data$TARGET_WINS)) + 
  geom_point() + geom_smooth(method=lm) + labs(x="TEAM_BASERUN_CS", y="TARGET_WINS", title="")
```

```{r}
summary(lm(TARGET_WINS ~ TEAM_BASERUN_CS, data=data))
```

---

## FIELDING VARIABLES 


### TEAM_FIELDING_E

Errors.

No missing values. 

Negative relationship (expected theoretical effect is negative). No obvious curved patterns observed from visually inspecting the scatter plots only.

Linear regression predictor coefficient is -0.012205 (points contributed to TARGET_WINS with every unit increase). The p-value is <2e-16 *** (predictor is significant). The R-squared is 0.03115 (explains about 3.1% variability of response variable).


```{r echo=FALSE, warning=FALSE}
ggplot(data, aes(x=data$TEAM_FIELDING_E, y=data$TARGET_WINS)) + 
  geom_point() + geom_smooth(method=lm) + labs(x="TEAM_FIELDING_E", y="TARGET_WINS", title="")
```

```{r}
summary(lm(TARGET_WINS ~ TEAM_FIELDING_E, data=data))
```

---

### TEAM_FIELDING_DP (286 missing)

Double plays. 

286 missing values. Results of simple linear regression below suggests that this variable is not significant. 

Recommendation is to drop this variable from the model. 

Negative relationship (expected theoretical effect is positive). No obvious curved patterns observed from visually inspecting the scatter plots only.

Linear regression predictor coefficient is -0.01853 (points contributed to TARGET_WINS with every unit increase). The p-value is 0.12 (predictor is not significant at 5% level). The R-squared is 0.001215 (explains less than 1% variability of response variable).


```{r echo=FALSE, warning=FALSE}
ggplot(data, aes(x=data$TEAM_FIELDING_DP, y=data$TARGET_WINS)) + 
  geom_point() + geom_smooth(method=lm) + labs(x="TEAM_FIELDING_DP", y="TARGET_WINS", title="")
```

```{r}
summary(lm(TARGET_WINS ~ TEAM_FIELDING_DP, data=data))
```


