---
title: "DATA 621---Assignment no. 2"
author: "Critical Thinking Group 2"
date: "October 10, 2019"
output:
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE)

# Load libraries
library(caret)
library(pROC)
```


# Executive Overview

We explore metrics to evaluate the performance of classification algorithms, including custom R functions, on a real dataset:

* Accuracy
* Classification error rate
* Precision
* Sensitivity
* F1 score
* ROC curve

We also examine the functionality of the `caret` and `pROC` packages.

```{r, echo=FALSE}
# load data
df <- read.csv('classification-output-data.csv', stringsAsFactors=FALSE)
```



# 2. Confusion matrix







# 3. Accuracy







# 4. Classification error rate







# 5. Precision

[BEN]




# 6. Sensitivity

[BEN]





# 7. Specificity







# 8. F1 score

[BEN]





# 9. F1 score: Bounds







# 10. ROC curve







# 11. In Practice







# 12. `caret` package







# 13. `pROC` package







